Approximation as Differentiation

### 2025-05-05 appendix/notes/auto-diff.tex
Clarify that forward/backward derivatives are determined by f and given by the Jacobean matrix of f at x and its transpose.

Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
(tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in \RR^m$
is a linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$. Because $\RR^n$ is Euclidean (flat),
$\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually $\pushf{f}_x: \RR^m \linearto \RR^n$. The
\emph{backward derivative} (cotangent map or pullback) $\pullf{f}_x$  is a linear map $\RR^n \linearto \RR^m$.

>>>>

Write $\tangents_x(\RR^n)$ for the tangent space at a point $x \in \RR^n$. Then the \emph{forward derivative}
(tangent map or pushforward) $\pushf{f}_x$ of a differentiable function $f: \RR^m \to \RR^n$ at $x \in \RR^m$
is the unique linear map $\tangents_x(\RR^m) \linearto \tangents_{f(x)}(\RR^n)$ given by the Jacobean matrix
of $f$ at $x$. Since $\RR^n$ is flat, $\tangents_x(\RR^n)$ is naturally isomorphic to $\RR^n$, so actually
$\pushf{f}_x$ is a linear map $\RR^m \linearto \RR^n$. The \emph{backward derivative} (cotangent map or
pullback) $\pullf{f}_x$ is the unique linear map $\RR^n \linearto \RR^m$ given by the transpose (adjoint) of
the Jacobean matrix of $f$ at $x$.

### 2025-05-05 appendix/notes/grothendieck.tex
Spelling out auto-diff via Grothendieck.
Clarify that forward/backward derivatives are determined by f and given by the Jacobean matrix of f at x and its transpose.

\begin{definition}[Grothendieck construction for $F$]
Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction}
for $T$ is the category $\Grothendieck{}T$ which has:
\begin{itemize}
\item as objects, all pairs $(X, \partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of
$T(X)$;
\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to
Y$ in $\cat{C}$ and morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
\end{itemize}
\end{definition}

>>>>

\begin{definition}[Grothendieck construction for $F$]
\label{def:grothendieck:Grothendieck}
Suppose an indexed category $T: \cat{C}^\op \to \Cat$. The \emph{Grothendieck construction}
for $T$ is the category $\Grothendieck{}T$ which has:
\begin{itemize}
\item as objects, all pairs $(X, \partial X)$ of an object $X$ of $\cat{C}$ and an object $\partial X$ of
$T(X)$;
\item as morphisms $(X, \partial X) \to (Y, \partial Y)$, all pairs $(f, \partial f)$ of morphisms $f: X \to
Y$ in $\cat{C}$ and morphisms $\partial f: \partial X \to T(f)(\partial Y)$ in $T(X)$.
\end{itemize}
\end{definition}

\noindent The $\partial X$ notation is justified by the fact that the Grothendieck construction formalises the
idea of tangent spaces and derivatives.
